---
title: "Predective_Analytics_Inclass"
author: "Swe Zin Ye"
date: "2025-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Using Boston hosuign price - Make house price prediction
Step invlved 
1. Read the data
2. Clean the data 
3. Correlation analysis 
4. Train nd test split 
5. create a model 
6. Make predictions 
7. Capture train and test errors 

```{r}

library(ggcorrplot)
library(ggplot2)
library(caret)

```




```{r cars}

house_prices <-read.csv("06_Regression_Boston_Housing_Prices.csv")

```


```{r}
#step 2 Clean data 
summary(house_prices)
# clean the data just by replacing the NA value with mean values 

#summary(replace(house_prices$CRIM, is.na(house_prices$CRIM),         #               mean(house_prices$CRIM,
  #      na.rm = TRUE)))

house_prices_clean <- data.frame(
  lapply(house_prices, FUN = function(x) {
    replace(x, is.na(x), 
        mean(x, na.rm = TRUE))
  })
)
summary(house_prices_clean)

```


```{r}
#step 3: Correlation Analysis 
# Cor function ( correlation)
# - scatter plot ( Matrix of plots) pairs()
#correlation matrix an numbers using ggcorrplot 

ggcorrplot(corr= cor(house_prices_clean),
           lab=TRUE,
           lab_size = 2,
           type = "lower")

  



```



```{r}
#Step 4: split the data randomly into Traiing ( 70 % to 80%) and Testing ( 30 % to 20%) 

set.seed(1234)
sample_postions<- sample(c("for_training","for_testing"),
                         size = nrow(house_prices_clean),
                         prob = c(0.8,0.2),
                         replace = TRUE)
head(sample_postions, 10)
                        

```


```{r}
#step4 (Cont): Create Training_Sample and Testing_Sample using Sample_ Positions 
Training_Sample <- house_prices_clean[sample_postions == "for_training",]
Testing_Sample <- house_prices_clean[sample_postions == "for_testing",]

```



```{r}
#Step 5: create a model using training sample 
#1. step forward/backward 
#2. Lasso
#3.Ridge

model <- lm(MEDV ~.,data = Training_Sample)
summary(model)


```

```{r}
#Setp 6: capture the training error (MSE)-. degree of freedom 
Training_Error <- sum((model$residuals)^2)/nrow(Traing_Sample)
Training_Error

```

```{r}

#Step 7: Make predections 
predictions <- predict(model,Testing_Sample[c(-14)]) # remove the predected model colum , chek data set and remove the column number e.g price of house 
Testing_Sample$Predicted_Output <- predictions
testing_Error <-Testing_Sample$MEDV - Testing_Sample$Predicted_Output
Testing_Error <- sum(testing_Error^2)/length(testing_Error)
Testing_Error


```
```{r}

#step-wise regression implementation 
#AIC value is lesser , better 
minimal_model <-lm(MEDV ~ 1, data = Training_Sample)
final_model <- step(minimal_model,
                    scope = ~ CRIM + ZN + INDUS + CHAS + NOX + RM + AGE + DIS + RAD + TAX + PTRATIO + B + LSTAT)
summary(final_model)

```



```{r}
# Step : Make predictions and calculate test MSE 

Training_Error <- sum((final_model$residuals)^2)/nrow(Traing_Sample)
Training_Error

```


```{r}
#Step 7: Make predections 
predictions <- predict(final_model,Testing_Sample[c(-14)]) # remove the predected model colum , chek data set and remove the column number e.g price of house 
Testing_Sample$Predicted_Output <- predictions
testing_Error <-Testing_Sample$MEDV - Testing_Sample$Predicted_Output
Testing_Error <- sum(testing_Error^2)/length(testing_Error)
Testing_Error


```


```{r}

#Run Ridge Regression 

training_parameters <- trainControl(method = "repeatedcv",
                                    number = 10, 
                                    repeats = 5,
                                    verboseIter = F)


ridge_model <- train(MEDV ~., 
                data =  Training_Sample,
                method = 'glmnet',
                tuneGrid = expand.grid(alpha = 0, 
                                       lambda = seq(0, 5, length = 50)), # set alpha = 0 , lambda = 0 to 5 and set of 50 
                trControl = training_parameters)

ridge_model$bestTune




```



```{r}
#Make prediction using ridge 

predictions2 <- predict(ridge_model,Testing_Sample)
Testing_Sample$Predicted_Output2 <- predictions2

Testing_Error2 <- sum((Testing_Sample$MEDV - Testing_Sample$Predicted_Output2)^2)/95

Testing_Error2


```



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
